# This is the code used to generate Hannah's NMDS plots.
# https://www.rpubs.com/RGrieger/545184

# It doesn't seem like any preprocessing or transformation was performed.

# Original matrix has ASVs as row labels and sample names as column headers. 
counts_tab <- read.csv("ASVs_counts_Mar2023.tsv", header=T, row.names=1 ,check.names=F, sep="\t")
counts_tab <- counts_tab %>% select(-c('HDK-DNAexNegLot169030916-30cyc', 'HDK22-KML-sand-dry','HDK22-KML-sand-wet', 'HDK-MAR-PCR-BLANK'))


# Create dataframe of ASV's and sample names in order needed for vegan package.
# The t function is a matrix transpose (NOT a transformation). This will make it so that ASVs are columns and samples are rows.
data1 <- data.frame(t(counts_tab[]))

# Load in the metadata.
sample_info_tab <- read.table("DNA_input_Mar2023.txt", sep = "\t",  colClasses = "character", header = TRUE)

# Convert sample info and environemtal variables into format needed for vegan.
samp1 <- sample_info_tab[,-1]
rownames(samp1) <- sample_info_tab[,1]


# Removing unneeded columns for ordination fitting -- all columns except for the environmental data to test to see if it correlates with the NMDS plot.
# samp1 <- samp1 %>% select(-c(Serial.code))
# Change values to numeric value type.
samp1$Flux = as.numeric(samp1$Flux)


# Create NMDS using the distance formula (in this case, bray-curtis)
# Check stress in output.
# Default MDS engine is monoMDS. Also wascores=True and expand=True... maybe this is what I am missing?
# wascores doesn't affect the actual ordination, but populates a "wa" attribute in the output. Is this used for plotting?
stations.mds <- metaMDS(data1, distance = "bray", atuotransform = FALSE)


#create envfiorntmal vectors fit OPTIONAL
stations.envfit <- envfit(stations.mds, samp1, permutations = 999, na.rm = TRUE)

#create species fit - will run for a long time OPTIONAL
#stations.spp.fit <- envfit(stations.mds, asv_df, permutations = 999)


#save
df_env <- data.frame((stations.envfit$vectors)$arrows, (stations.envfit$vectors)$r, (stations.envfit$vectors)$pvals)

#df_spp <- data.frame((stations.spp.fit$vectors)$arrows, (stations.spp.fit$vectors)$r, (stations.spp.fit$vectors)$pvals)

#write.csv(df_env, "stations_envfit_sept2023.csv")
#write.csv(df_spp, "stations_spp_sept2023.csv")


```





#### PLOT VEGAN NMDS - environmetal variables
```{r eval=F}
#basic plot
plot(stations.mds)

#To plot the output from the mds using ggplot a new datasheet needs to be created which contains the x,y points for each site. You can do this by calling the scores of you mds.

site.scrs <- as.data.frame(scores(stations.mds, display = "sites")) #save NMDS results into dataframe

site.scrs <- cbind(site.scrs, Site = samp1$Site) #add grouping variable "station" to dataframe

site.scrs <- cbind(site.scrs, Soil.depth = samp1$Soil.depth)

site.scrs <- cbind(site.scrs, Season = samp1$Season)

site.scrs <- cbind(site.scrs, Flux = samp1$Flux)

site.scrs <- cbind(site.scrs, Serial.code = samp1$Serial.code)


head(site.scrs)


```


To show environmental extrinsic variables another datasheet needs to be created

```{r eval=F}
env.scores.stations <- as.data.frame(scores(stations.envfit, display = "vectors")) #extracts relevant scores from envifit

env.scores.stations <- cbind(env.scores.stations, env.variables = rownames(env.scores.stations)) #and then gives them their names

env.scores.stations <- cbind(env.scores.stations, pval = stations.envfit$vectors$pvals) # add pvalues to dataframe

sig.env.scrs <- subset(env.scores.stations, pval<=0.05) #subset data to show variables significant at 0.05

env.scores.stations


```


#### PLOT
```{r eval=F}

nmds.plot.stations <- ggplot(site.scrs, aes(x=NMDS1, y=NMDS2))+ #sets up the plot
  geom_point(aes(NMDS1, NMDS2, size = Flux, color = Site, label = Serial.code))+ #adds site points to plot
  coord_fixed()+
  theme_classic()+ 
  theme(panel.background = element_rect(fill = NA, colour = "black", size = 1, linetype = "solid"))+
  labs(colour = "station", shape = "locale")+ # add legend labels for Management and Landuse
  theme(legend.position = "right", legend.text = element_text(size = 12), legend.title = element_text(size = 12), axis.text = element_text(size = 10)) # add legend at right of plot

nmds.plot.stations + labs(title = "Basic ordination plot") #displays plot

to_export <- nmds.plot.stations + labs(title = "Basic ordination plot") 

ggsave("NMDS_1.pdf", to_export)

ggplotly(nmds.plot.stations + labs(title = "Basic ordination plot"))


nmds.plot.stations <- ggplot(site.scrs, aes(x=NMDS1, y=NMDS2))+ #sets up the plot
  geom_point(aes(NMDS1, NMDS2, size = Flux, color = Site, shape = Soil.depth  ))+ #adds site points to plot
  coord_fixed()+
  theme_classic()+ 
  theme(panel.background = element_rect(fill = NA, colour = "black", size = 1, linetype = "solid"))+
  labs(colour = "station", shape = "locale")+ # add legend labels for Management and Landuse
  theme(legend.position = "right", legend.text = element_text(size = 12), legend.title = element_text(size = 12), axis.text = element_text(size = 10)) # add legend at right of plot

nmds.plot.stations + labs(title = "Basic ordination plot") #displays plot


to_export2 <- nmds.plot.stations + labs(title = "Basic ordination plot") 

ggsave("NMDS_2.pdf", to_export2)


#Significant Environmental Variables

nmds.plot.stations+
  geom_segment(data = sig.env.scrs, aes(x = 0, xend=NMDS1, y=0, yend=NMDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd=0.3) + #add vector arrows of significant env variables
  ggrepel::geom_text_repel(data = sig.env.scrs, aes(x=NMDS1, y=NMDS2, label = env.variables), cex = 4, direction = "both", segment.size = 0.25)+ #add labels for env variables
  labs(title="Ordination with environmental vectors")


to_save <- nmds.plot.stations_only_samples_2+
  geom_segment(data = sig.env.scrs, aes(x = 0, xend=NMDS1, y=0, yend=NMDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd=0.3) + #add vector arrows of significant env variables
 # ggrepel::geom_text_repel(data = sig.env.scrs, aes(x=NMDS1, y=NMDS2, label = env.variables), cex = 4, direction = "both", segment.size = 0.25)+ #add labels for env variables
  labs(title="Ordination with environmental vectors")

to_save

ggsave("NMDS_samplesonly.png", to_save)



```


#CONTOUR LINES ON NMDS
```{r eval=F}

 plot(stations.mds, type = "n")  
points(stations.mds, display = "sites") 
ordisurf(stations.mds,samp1$Flux, add = TRUE)




pdf(file = "FluxPlot.pdf",   # The directory you want to save the file in
    width = 4, # The width of the plot in inches
    height = 4) # The height of the plot in inches

 plot(stations.mds, type = "n")  
points(stations.mds, display = "sites") 
ordisurf(stations.mds,samp1$Flux, add = TRUE)


# Step 3: Run dev.off() to create the file!
dev.off()

```




```{r eval=F}

plot(stations.mds, display = "sites")

#pl <- with(stations.mds) ordiellipse(stations.mds, site.scrs$station, kind="se", conf=0.95, lwd=2, draw = "polygon", col=1:4, border=1:4, alpha=63)

ordiellipse(stations.mds, site.scrs$station, kind="se", conf=0.99, lwd=2, draw = "polygon")


```



